---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rademacher

<!-- badges: start -->
<!-- badges: end -->

The goal of `rademacher` is to help generate rademacher random variables as quickly as possible using low-level C++ and bit operations. Note that this function is now implemented in the `dqrng` and is even faster than this version (see benchmark below). Use that if you don't mind adding dependencies to your package (a few C++ libraries are used in dqrng).

The Rademacher distribution draws -1 and 1 with probability 0.5 each. The way I generate $N$ is (approximately) as follows: 

1. Draw `N / 31` integers from 1:(2^31 - 1).
2. Each integer has 31 bits (0/1) variables and each bit is randomly generated. I grab all the bits and do $b * 2 - 1$ to convert to $-1, 1$.

This is totally open source. You can copy and past to your package without any credit needed. 

There is a developmental 64-bit version, but I have not figured out a good way to handle seeds, yet (currently just requires a seed be given). It was primarily written for testing purposes, but if you want the fastest version, use the package `dqrng`.

## Installation

You can install the development version of rademacher like so:

``` r
devtools::install_github("kylebutts/rademacher")
```


## Benchmark

The fastest way to generate using base R is using `runif()`, so I will compare to that. I also add in `dqrng::dqrrademacher()` that does the same thing as this package but with a faster RNG. 

```{r benchmark}
library(rademacher)
library(dqrng)
library(bench)

runif_rademacher <- function(n) { 
  (runif(n) > 0.5) * 2 - 1
}

n = 1e7
bench::mark(
  runif_rademacher  = runif_rademacher(n),
  sample_rademacher = rademacher::sample_rademacher(n),
  dqrrademacher     = dqrng::dqrrademacher(n),
  check = FALSE
)

n = 1000
bench::mark(
  runif_rademacher  = runif_rademacher(n),
  sample_rademacher = rademacher::sample_rademacher(n),
  dqrrademacher     = dqrng::dqrrademacher(n),
  check = FALSE
)
```

This is totally open source (CC0 license). You can copy and paste to your package without any credit needed. 
